---
title: "STAT X12 Project"
output:
  word_document:
    fig_height: 5
    fig_width: 8
date: ""
author: Name
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
options(show.signif.stars = FALSE)

library(ggplot2)
library(ggthemes)
library(tidyverse)
library(car)
library(effects)
library(janitor)
library(readxl)
library(catstats2)
library(mosaic)
theme_set(theme_bw()) #Prevents need for + theme_bw() in ggplots
```


```{r}
library(mi)
library(haven)
mydata <- read_sav("C:\\Users\\nitas\\Downloads\\Overreportinginhandwashingselfreports_Final.sav")
```

```{r}
my_data_m <- missing_data.frame(data.frame(mydata))
image(my_data_m)
colnames(my_data_m)
dim(my_data_m)
```


```{r}
colSums(!is.na(mydata))
```
```{r}
colnames(mydata)[which(names(mydata) == "OR_SR_food")] <- "OvR_F"
colnames(mydata)[which(names(mydata) == "OR_SR_stool")] <- "OvR_S"
colnames(mydata)[which(names(mydata) == "E315_327_SD")] <- "MCSCS"
colnames(mydata)[which(names(mydata) == "E304_01")] <- "Need_Confirm"
colnames(mydata)[which(names(mydata) == "E305_01")] <- "Group_Attach"
colnames(mydata)[which(names(mydata) == "E6134_01")] <- "Des_Food_N"
colnames(mydata)[which(names(mydata) == "E6156_01")] <- "Inj_Food_N"
colnames(mydata)[which(names(mydata) == "E5123_01")] <- "Des_Stool_N"
colnames(mydata)[which(names(mydata) == "E5145_01")] <- "Inj_Stool_N"
colnames(mydata)[which(names(mydata) == "E121.1_exHusb_cat")] <- "Pres_Oth_Adult"
colnames(mydata)[which(names(mydata) == "E341.1")] <- "Pres_Spouce"
colnames(mydata)[which(names(mydata) == "E301")] <- "Daily_Routine"
colnames(mydata)[which(names(mydata) == "E335_339_2_MT_01")] <- "Fr_Task_Tnt"
colnames(mydata)[which(names(mydata) == "E342")] <- "Est_Tendency"
colnames(mydata)[which(names(mydata) == "E713_HW_stool_01")] <- "HW_Knowl_Stool"
colnames(mydata)[which(names(mydata) == "E713_HW_food_01")] <- "HW_Knowl_Food"
colnames(mydata)[which(names(mydata) == "E328_E334_DISS_Rationalisation_01")] <- "Rationalisation"
colnames(mydata)[which(names(mydata) == "E713_stoolxE32834")] <- "Interaction_KW_R_Stool"
colnames(mydata)[which(names(mydata) == "E713_foodxE32834")] <- "Interaction_KW_R_Food"
colnames(mydata)[which(names(mydata) == "E306_repoled_01")] <- "Loaded_Que_Word"
colnames(mydata)[which(names(mydata) == "MeanHW_01")] <- "SR_Food_Stool"
colnames(mydata)[which(names(mydata) == "Version_numerisch")] <- "Sc_Que_Version"
colnames(mydata)[which(names(mydata) == "E306_E307_FW_umgepolt_01")] <- "Sc_Forgiving"
colnames(mydata)[which(names(mydata) == "E309_E311_E306_context_01_repol")] <- "Sc_Que_Context"
colnames(mydata)[which(names(mydata) == "OR_ICR_stool")] <- "OR_Script_Stool"
colnames(mydata)[which(names(mydata) == "OR_ICR_food")] <- "OR_Script_Food"
 

```


Subset for food related Data
```{r}
mydata2 <- mydata[ , c(35:ncol(mydata))]
mydata2 <- mydata2[ , c(-5,-7,-14,-17,-24)]


my_data2_m <- missing_data.frame(data.frame(mydata2))
image(my_data2_m)
colnames(my_data2_m)
dim(my_data2_m)
```
Removing the script related variables from Food related dataset.
```{r}
mydata2_a <- mydata2[ , c(-15,-16,-17,-18,-19,-20)]


my_data2_m <- missing_data.frame(data.frame(mydata2_a))
image(my_data2_m)
dim(my_data2_m)

mydata2_b <- mydata2_a %>% drop_na()
my_data2b_m <- missing_data.frame(data.frame(mydata2_b))
image(my_data2b_m)
dim(my_data2b_m)
```      

Correlation to find the answers of first 9 hypothesis.(Food)
```{r fig.height=6,fig.width=10}
library(psych)
pairs.panels(mydata2_b,smooth=FALSE,ellipses=FALSE,digits=4)
{pairs.panels}
```

Fitting the Linear Models
```{r}

lm_F_S_R = lm(OvR_F~MCSCS + Need_Confirm  + Des_Food_N + Inj_Food_N + Fr_Task_Tnt  +Est_Tendency+ HW_Knowl_Food + Rationalisation,  data=mydata2_b) 
vif(lm_F_S_R)
summary(lm_F_S_R)

```


Checking the Assumptions for the linear Model
```{r fig.height=8,fig.width=10}

library(ggResidpanel)
library(haven)
mydata2_b <- zap_formats(zap_labels(mydata2_b))
plot(allEffects(lm_F_S_R, residuals = T), grid = T)
resid_xpanel(lm_F_S_R, yvar = "response")
resid_panel(lm_F_S_R, "R", alpha=0.1)

```

From here we will do our own analysis on the 

Research Question:
We will systematically investigate various model combinations to identify potential predictors that may offer improved insights into the over reporting of hand washing behavior."

From here on we will only work for the Food related over reporting of hand wash.
Lets pick the full model according to paper. 

```{r}
options(na.action = "na.fail")
library(MuMIn) #Load the multi-model inference package
# prevent fitting sub-models to different datasets
options(na.action = "na.fail") #Must be run to use dredge
res_lm_F_S_R <- dredge(lm_F_S_R, rank = "AIC", extra = "R^2") 
subset(res_lm_F_S_R, delta<6)
dim(res_lm_F_S_R)

```


```{r}

MMFood <- lm(OvR_F ~1,data=mydata2_b)
AIC(MMFood)
head(res_lm_F_S_R,3)
```
There were 256 models explored. The top selected model and mean only model has AIC difference of 99.82. which provides us strong evidence for our top model.
Top model with AIC 182, contains Des_Fod_N , Fr_Tsk_Tnt ,HW_Knw_Fod ,Inj_Fod_N and rationalization as predictors 


Next we tested interaction 
```{r fig.height=8,fig.width=10}

lm_F_S_R_2 = lm(OvR_F~ (Des_Food_N + Inj_Food_N + Fr_Task_Tnt  + Rationalisation)^2,  data=mydata2_b) 
res_lm_F_S_R_2 <- dredge(lm_F_S_R_2, rank = "AIC", extra = "R^2") 
head(subset(res_lm_F_S_R_2, delta<6),3)
dim(res_lm_F_S_R_2)
```


Effect Plot
```{r}
resid_panel(model = lm_F_S_R, plot= 'R', alpha=0.1, smoother= T)
resid_panel(model = lm_F_S_R_2, plot= 'R', alpha=0.1, smoother= T)
```
```{r catRmd,asis=TRUE}
rmd<-read_lines("STAT512ProjectFinal.Rmd")
cat(paste(rmd,"\n"))
```